---
title: "Portfolio"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 2
    theme: lumen
    highlight: tango
---

```{r general-options-document-format, include=FALSE}
knitr::opts_chunk$set(tidy = 'styler', warning = F, message = F, echo = F, error = F)  # Set knitting options general
options(tinytex.verbose = T, dplyr.summarise.inform = F)  # Debugging & Suppress summary info
```



Here, I am showcasing some selected Data Science & AI projects which I had the pleasure to work on in the last couple of years. 



# Automated Fossil Fuel Policy Tracker

As part of my Data Science Fellowship with <a href="https://faculty.ai/" target="_blank">Faculty AI</a>, I worked together with the *The Fossil Fuel Non-Proliferation Treaty Initiative (FFNPT)*. The <a href="https://fossilfueltreaty.org/" target="_blank">FFNPT Initiative</a> is a global NGO empowering policy advisers, activists, and researchers to facilitate a rapid global transition away from fossil fuel toward clean energy. 

The FFNPT Initiative has recently been endorsed by <a href="https://www.theguardian.com/environment/2021/apr/21/101-nobel-laureates-call-for-global-fossil-fuel-non-proliferation-treaty" target="_blank">101 Nobel Laureates</a>  and is thus rapidly gaining international traction.

I had the pleasure to work with the FFNPT on a Data Science project in which I programmed an interactive outline app, using web-scraping and Natural Language Processing (Artificial Neural Network) to automatically identify, filter, and visualize relevant policies globally.

 * <a href="https://fossil-fuel-tracker.herokuapp.com/" target="_blank">FFNPT App</a>
 
In the video below I present the app in more detail.

```{r echo=F,include=F,error=F}
library(vembedr)
```


```{r echo=F,error=F}
embed_youtube("aCP8X7Pnh4o", width = NULL,
  height = 500,
  ratio = c("16by9", "4by3"),
  frameborder = 0,
  allowfullscreen = TRUE) %>% 
  use_align(align = "center")
```


# Climate Change and Machine Learning

In my role as a Postdoctoral Research Associate & Data Scientist with <a href="https://www.sdmlab.psychol.cam.ac.uk/staff/dr-fatih-uenal" target="_blank">The University of Cambridge</a>, I have worked with big data sets using Machine Learning Algorithms to identify the most robust predictors of Climate Change Attitudes and Policy Preferences. Some of the results of this line of work is available 

 * <a href="https://fuenal.github.io/dataism/ClimateChangeAI.html" target="_blank">Climate Change Perceptions: A Machine Learning Approach</a> 



---


