---
title: "Syllabus"
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 4
---

### Data Science for Social Sciences (Dataism 101)

### Instructor Info

Lectures

- Time:	TBD 	Room: TBD

Instructor

- Dr. Fatih Uenal	Old Cavendish Labarotory
- Office hours:	by appointment
- fu212@cam.ac.uk

---

### Lecture Schedule

```{r, message=FALSE, warning=FALSE, echo=FALSE}
#library(data.table)
library(readxl)
Lectures <- read_excel("LectureSpring2019.xlsx", sheet = 1, col_names = TRUE, col_types = NULL, na = "", skip = 0)
Lectures[is.na(Lectures)]<- " "
#Lectures <- fread("Lectures.csv")
knitr::kable(Lectures)

```


Website:  [https://fuenal.github.io/dataism/](https://fuenal.github.io/dataism/)

---

### Catalog Description

Basic descriptive and inferential statistics including the elements of experimental designs in social sciences research. Additional Data Science Techniques and advanced statistical analyses (optional)

### Course Objectives

“Data science” is a catch-all term used to describe the practice of working with and analyzing messy data sources to draw meaningful conclusions using techniques developed by computer scientists and computational statisticians. The purpose of this course is to give students who are training as quantitative social scientists a broad introduction to this skillset via the statistical programming language, R. You will learn how to conduct many statistical analyses such as univariate statistics (e.g., ANOVA, correlation, regression) in R that you have already done in SPSS, Excel, or another such program. Additionally, we will build on this foundation to explore new skillsets uncommon in the social sciences, such as natural language processing, automated data curation, and machine learning. I have also included more advanced statistical analysis (SEM) as optional choice.

At the end of this course you will be able to:

1. To answer research questions in Social Sciences (e.g., Psychology) with data

2. Understand the basics of research designs in Social Sciences, and how they relate to data-analysis strategies

3. Develop an intuitive, practical, and conceptual understanding of strategies for asking and answering questions with data

4. To use R, a free and open-source statistics software program.

5. Develop a basic understanding of frequently used Data Science Techniques.

6. Practice your newly acquired skills with interesting, interactive and fun porjects.

NOTE: The schedule and procedures described in this syllabus are subject to change depending on specific needs and requirements. You will always be notified of changes on the homepage (see "last update"). 

---

### Resources (freely available)

The reading materials for this course are selected from freely available and frequently used sources online as web-books. Each of these sources have their own strengths and merits. In curating this course, I have tried to create the best fit between reading material and online courses (Datacamp). I have also tried to present a balanced mix between Data Science and classical Social Science related Intstruction books. All materials linked in this course include further relevant material on a broad range of relevant topic and I encourage you to browse through them. Also, I encourage you to make use of the existing exercises within each of the here referenced resources: 

1.	Neth, H. (2020).  Data Science for Psychologists.  Social Psychology and Decision Sciences, University of Konstanz, Germany.  Textbook and R package ds4psy (version 0.2.1, May 6, 2020).  eBook available at https://bookdown.org/hneth/ds4psy/;  R package at https://CRAN.R-project.org/package=ds4psy

2.	Navarro, D. (2019). Learning statistics with R: A tutorial for psychology students and other beginners. (Version 0.6.1, June 19, 2019). Danielle Navarro (bookdown translation: Emily Kothe) eBook available at https://learningstatisticswithr.com/book/

3.	Irizarry, R. (2019). Introduction to Data Science. Data Analysis and Prediction Algorithms with R. eBook available at https://rafalab.github.io/dsbook/

4.	Wickham, H. & Grolemund, G. (2017). R for Data Science Import, Tidy, Transform, Visualize, and Model Data. eBook available at https://r4ds.had.co.nz/ 

5.	Field, A., Miles, J., & Field, Z. (2012). Discovering Statistics Using R. London, UK: Sage Publications. 

6.	Silge, J. & Robinson, D. (2017). Text Mining with R: A Tidy Approach. eBook available at https://www.tidytextmining.com/

7.	Harrer, M., Cuijpers, P., Furukawa, T.A, & Ebert, D. D. (2019). Doing Meta-Analysis in R: A Hands-on Guide. eBook available at https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/. 

---

### Additional resources

Supplementary material will be posted on the course website. For example, check out the [resources page](https://fuenal.github.io/dataism/Resources.html) on this website.

---

### Course Structure:

#### The Lecture

There will be two, 1 hour 15 minute, lectures each week. The lecture covers material that is available in the text book. The course schedule outlines chapters that **should be read prior to each lecture**. It is very important to read the textbook before you come to lecture. I've said it twice, now again. **Read the assigned parts of the textbook before each lecture**. Read it at home, read it on the bus, read it in a tree, read it at the library. Just read it.

You will be tested on material from the lectures and textbook in two exams: a midterm, and a cumulative final exam.

#### The Lab

There will be one two hour lab each week (1 hour 50 minutes). Every week you will be completing a structured set of activities that we have planned for you in the [lab manual](https://crumplab.github.io/statisticsLab/). This is where you will spend time learning how to use computer software, in particular R, and R-studio, to do statistics on real data. 

---

### Grading (total of 100)

Your final grade combines your performance in the lab and lecture portions of the course. The percentage grade is determined from the following components.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(data.table)
grading <- fread("Grading.csv")
knitr::kable(grading)

```

Percentage grades are converted to letter grades according to the following rubric.

|Letter grade|Range|
|:--|:--|
|F	|0-59.99|
|D-	|60-63.32|
|D	|63.33-66.66|
|D+	|66.67-69.99|
|C-	|70-72.99|
|C	|73-76.66|
|C+	|76.67-79.99|
|B-	|80-83.32|
|B	|83.33-86.66|
|B+	|86.67-89.99|
|A-	|90-93.32|
|A	|93.33-96.66|
|A+	|96.67-100|

The course allows for some flexibility in determining your final grade. If you perform better on the final exam than on the midterm, you may replace the midterm grade with your higher final grade (the reverse does not apply).  

The instructor reserves the right to determine whether the final grade depends more heavily on a scale or a curve.

---

#### Lab Work (40%)


As you can see, your lab work takes up 40% of your entire grade. Compared to everything else, this is the single biggest part of your entire grade. There are 11 graded labs.

The lab work for each week is split up into three parts. Your grade for each lab will be broken down as follows (4 total points for each lab). 

The whole point of the labs is **to train you how to do things**. You will notice that half of the lab is pass/fail, you just have to show up and follow along, and hand in your work to get half of the grade. The third part is graded, but you will see why in a moment, we're giving you practice for the kinds of the questions you will see on the midterm and final.

**Part 1. Learn** 1 point

- Part 1 grading is pass/fail:
    
    a. Pass =  you must attend lab, and hand in your assignment with a completed part 1
    b. Fail = you don't attend lab; or you do, but you don't hand in your assignment.
        
- Part 1 activities: You will follow the step-by-step instructions in the lab manual to complete data analysis tasks using R. Don't worry this is actually pretty easy, we are mainly asking you to copy/paste things and press enter.

**Part 2. Generalize** 1 point

- Part 2 grading is pass/fail:
    
    a. Pass =  you must attend lab, and hand in your assignment with your part 2 attempt
    b. Fail = you don't attend lab; or you do, but you don't hand in your assignment.

- Part 2 activities: After you see how we did something in Part 1, you try to do something very similar in part 2. You will show your work in a file that we give you. You may or may not be able to successfully complete the generalization assignment. That is OK. You will pa

**Part 3. Think and write** 2 points

The third part of each lab will have some questions that require written answers. Your answers to these questions will be graded by your TA, who will give you feedback on your answers.

Each question will come with a clear rubric about how the question will be graded.

<!--
**Group work** is possible for part 3. If you want to form a group other students for part 3, then you can do this. You can work together to come up with a written answer to the question(s) in part 3. If you write as a group, then you MUST hand-in as a group. That means everybody in the group hands in the "exact" same written work to the question, and everybody in the group will get the exact same grade for the question.

**IMPORTANT: Part 3 of the labs will prepare you for the Midterm and Final**. Half of the midterm and half of the final involve writing. In fact, some of the long answer questions on the midterm and final will be taken from your labs, so you will already have had practice thinking about the questions. However, the **midterm and final** are **individual** assignments, so if you do group work in lab, make sure you are prepared to give answers individually for the midterm and final.

-->

---

#### Weekly Quizzes (25%)

Each week you will be given an online quiz on Blackboard. It will typically be around 10 questions about the material for that week. Here are some things you should know about the quizzes:

1. You will have the entire week to complete the quiz, and **you will be able to take the quiz over and over as many times as you want.**

2. Your grade on the quiz will always be taken from the last attempt.

3. Many of the quiz questions, or very similar ones, will be on the midterm and the final. The purpose of the quizzes is to give you practice with the kinds of questions that will be on the midterm and final. If you are wondering what kinds of questions are going to be on the midterm and final, **you will find out by taking the quizzes**.

---

#### Midterm (10%) and Final Exam (Cumulative 25%)

Everyone has questions about the midterm and final. Here are the answers to all of your questions. You already know the midterm is worth 20% of your final grade, and your final exam is worth 25% of the final grade.

**Q. What's on the exams?** A: the midterm and final have the same structure. Half of each exam is multiple choice questions, the other half is long-answer questions

**Q. What kind of multiple choice questions, I need examples...** A: All of the MC questions on the midterm and final will be like the ones in your weekly quizzes, which you can do over and over again. Yes, some of the midterm and final questions will be very similar (or even identical except for slightly different values) to the ones from the quizzes

**Q. What kind of long-answer questions, I need examples and feedback**: The long-answer questions will be very similar, and sometimes the same as the long-answers you write for Part 3 of each lab. You will get feedback from the lab instructor each week about your answers to those questions. In general you will have about 6-7 weeks of feedback on your answers before you take the exams. 

**Q. Is the final cumulative**? A: Yes, but it contains more things from the second half of the semester, roughly 75% second half things, and 25% first half things.

**Q. Is there some deal about the final and midterm, is this true?** A: Yes, if you perform better on the final exam than on the midterm, I will automatically replace your midterm grade with your higher final grade. Let's say you got 75% on the midterm and 85% on the final. I will then automatically give you 85% on both the midterm and final. However, the reverse does not apply.  If you got 85% on the midterm, and 75% on the final, you will receive those grades with no changes.


---

### Attendance

Lecture and Lab attendance is taken every lecture and lab.

**You will LOSE points on your quizzes and lab assignments for NOT ATTENDING LECTURE AND LAB**.

#### LECTURE ATTENDANCE POLICY

If a student is late or misses more than 3 classes, then each additional late/missed class will result in subtracting 1% from the 25% for the Weekly quizzes. We know that life happens and sometimes you can't make it to class. We are giving you you three classes where you can miss (because of things happening in your life that we don't need to know about) without having any impact on your grade (all lecture notes are online for missed classes, talk to a friend, read the textbook, etc.)

Lecture attendance is taken every lecture at the very beginning of class. Latecomers who miss attendance will not be allowed to record their attendance, and will be recorded as absent.

#### LAB ATTENDANCE POLICY

- Lab attendance is mandatory.
- Each student is given one unexcused lab absence with no penalty as long as your lab is submitted on time.  If you need to miss additional labs, please speak to your lab instructor and provide documentation regarding the reason you have to miss.
- If you miss more than one unexcused lab, you are still encouraged to hand your lab in and will receive up to 2 (rather than the usual 4) points.
- Labs for which no lab is handed in will receive a 0.


**Late Labs:**
Unless you have contacted your lab instructor regarding serious extenuating circumstances, there will be a penalty of .25 per day up to a total of 2 points for late labs.  This means that it is in your best interests to complete the lab, even late, as you can still receive up to 2 points (50% of the grade).  However, no lab reports will be accepted after three weeks beyond the due date.  Labs not submitted by three weeks past the due date will receive a zero.


---

### University’s policy on Academic Integrity

The faculty and administration of Brooklyn College support an environment free from cheating and plagiarism. As a student, **you are personally responsible for being aware of what constitutes cheating, and plagiarism; and, for avoiding both**. You can view the complete text of the CUNY Academic Integrity Policy here:
[http://www.brooklyn.cuny.edu/bc/policies](http://www.brooklyn.cuny.edu/bc/policies)
**If a faculty member suspects a violation of academic integrity, and upon investigation, confirms that violation, or if the student admits the violation, the faculty member MUST report the violation**.

---

### Missing an Exam or class

Make-up exams will be given only when there is a “good” reason for having missed an exam. For example, if a sudden illness causes you to miss an exam, then you should be prepared to provide me with a brief note signed by your physician. Contact me before an exam in the event that you anticipate missing one. In the event of an emergency, contact me as soon as possible. If you are missing a class for religious reasons refer to the state law regarding non-attendance because of religious beliefs (p. 64 in the Undergraduate Bulletin or p. 40 of the Graduate Bulletin).

WARNING: This is not the kind of course for which you do not need to come to class, and read the textbook at the last minute and expect to do well on tests and papers. The material in this course builds in complexity across classes, and doing well in this course will require a constant moderate level of effort. If you join the course later or are unsure about anything please ask.

---

### Accessibility

It is important to me that the course be accessible to all students. In order to receive disability-related academic accommodations students must first be registered with the Center for Student Disability Services. Students who have a documented disability or suspect they may have a disability are invited to set up an appointment with the Director of the Center for Student Disability Services, Ms. Valerie Stewart-Lovell at 718-951-5538. If you have already registered with the Center for Student Disability Services please provide me with the course accommodation form so we may discuss your specific accommodation. A guide and more information can be found here [http://catsweb.cuny.edu](http://catsweb.cuny.edu)

---

### Class Citizenship/Rules of Conduct

#### Cell phones

Your cell phone, PDA or other device must be turned off during class. If you are a habitual offender in this respect (i.e. it happens more than twice during the semester), you will be asked to leave the classroom. If you absolutely need to have your phone on during class—talk to me at the beginning of the semester.

#### Keep the room clean

Please pick up after yourself. Absolutely no food is allowed in the lecture and lab rooms.

--- 

### Email Correspondence

I will regularly use e-mail to send out announcements, changes in the syllabus, reminders about tests or due dates etc. It is your responsibility to check e-mail regularly to keep up-to-date with these announcements. I will use the e-mail address you have listed with the College. Therefore, please make sure that this is indeed the correct address.

#### General Help with Research and Writing

The Library maintains a collection of links to sites that can assist you with proper citation format and paraphrasing and quoting other authors at [Research & Writing Help](https://library.brooklyn.cuny.edu/resources/?service=research-help).  The Learning Center has writing tutors available to help you with your writing [http://lc.brooklyn.cuny.edu/](http://lc.brooklyn.cuny.edu/).

The best learning is done in conversation with others, whether they are people—classmates, teachers, friends—or texts—books, articles, essays, poems, films etc. It should not be a solitary process. However, the assignments that you hand in for this course must be done on your own, should represent your own thinking, and should be original work that you have done for this particular course. A good way to balance these two seemingly contradictory approaches (collaborative learning and original individually-produced work) without knowingly—or, even unwittingly—resorting to plagiarism or other forms of academic misconduct is to learn and meticulously observe the rules for citing the work of others (this could be the great point your roommate made that you used in your paper, it could be a well-turned phrase from an academic essay, or it could be anything in between). It is your responsibility to learn what constitutes plagiarism and the correct rules for citing sources—read the information on the following website carefully: [http://www.brooklyn.cuny.edu/bc/policies/](http://www.brooklyn.cuny.edu/bc/policies/). The bottom line is: passing off anyone’s words or ideas as your own for any reason whatsoever is plagiarism.
